---
title: "R Notebook"
output: html_notebook
author: Kevin Zarama
---
## Ejercicio cap 4

packages to open the file and read the file 
```{r}
install.packages("readxl")
library(readxl)
data_regmult <- read_excel("regmult.xls")
```
train the models with the required
```{r}
r1 <- lm(`I (T)` ~ `CE (T)` +  `CD (T)`, data_regmult)
r2 <- lm(`I (T)` ~ `Ldies (T)` + `LEl (T)`, data_regmult)
r3 <- lm(`I (T)` ~ `V (T)`, data_regmult)
r4 <- lm(`I (T)` ~ `CE (T)` + `LEl (T)`, data_regmult)
r5 <- lm(`I (T)` ~ `CD (T)` + `Ldies (T)`, data_regmult)
r6 <- lm(`I (T)` ~ `CE (T)` + `Ldies (T)` + `LEl (T)` + `V (T)`, data_regmult)
r7 <- lm(`I (T)` ~ `CE (T)` +  `CD (T)` + `V (T)`, data_regmult)
```
Add the results of the r squared in a vector to know the max
```{r}
which.max(c(summary(r1)$adj.r.squared,
summary(r2)$adj.r.squared,
summary(r3)$adj.r.squared,
summary(r4)$adj.r.squared,
summary(r5)$adj.r.squared,
summary(r6)$adj.r.squared,
summary(r7)$adj.r.squared))
```
Add the results of the aic in a vector to know the max
```{r}
which.max(c(AIC(r1),
AIC(r2),
AIC(r3),
AIC(r4),
AIC(r5),
AIC(r6),
AIC(r7)))
```
Add the results of the bic in a vector to know the max
```{r}
which.max(c(BIC(r1),
BIC(r2),
BIC(r3),
BIC(r4),
BIC(r5),
BIC(r6),
BIC(r7)))
```
According to r^2 the best model is the r6, while the two methods of the information criteria the best model is r3

Test of nested models
```{r}
anova(r1, r7)
```
Can´t reject the null hypothesis of that the restricted model r1 is better than the model not restricted r7, because the value of f is grater than p.
```{r}
anova(r2, r6)
```

```{r}
anova(r3, r6)
```

```{r}
anova(r3, r7)
```

```{r}
anova(r4, r6)
```